Low Complexity Adaptive Beamformer\\ for Active Sonar Imaging

### 2015-08-19 LCAforActiveSonarImaging/lca.tex
la2pdf commit - 2015:08:19	10:32:22	(Wed19.Aug)

The beamformer output $z[n]$ can now be expressed as the weighted sum of all the delayed data samples:
\begin{align}
z[n] = \w\H[n]\x[n] = \bmat{w_0[n]\\w_1[n]\\\vdots\\w_{M-1}[n]}^H \bmat{x_0[n]\\x_1[n]\\\vdots\\x_{M-1}[n]},\label{z}
\end{align}
where $w_m$ is the weight factor assigned to channel $m$. We will refer to a set of weights as a window, but it is also commonly called a taper function. Applying a window that trails off towards the edges improves noise suppression at the cost of reduced resolution. Real windows lead to symmetric responses, while complex weights allow asymmetric responses.

@@

MVDR seeks to minimize the power of the beamformer, under the constraint of a unity gain in the desired direction $\a$~\cite{Capon1969}:
%
\begin{align}
\underset{\w[n]}{\argmin}\, E\{|z[n]|^2\} &= \underset{\w[n]}{\argmin}\, \w[n]\x\x\H\w\H[n]\\
\text{subject to } \w[n]\a &= 1
\end{align}
%
where $\a$ is a steering vector. This is a convex optimization problem with the solution:
%
\begin{gather}
\vec w[n] = \frac{\Ri[n]\a}{\a\T\Ri[n]\a},\label{weights}
\end{gather}
where $\R=E\{\x[n]\x\H[n]\}\in\mathbb{C}^{M,M}$ is the spatial covariance matrix for the full array. The problem lies in estimating and inverting this spatial covariance matrix. Spatial averaging is needed to avoid signal cancellation~\cite{Kailath1985}, temporal averaging is needed to obtain true speckle statistics~\cite{Synnevag2009a}, and diagonal loading is needed to improve robustness to parameter errors~\cite{Cox1987,Maksym1979}. The steps are also needed to ensure that the covariance matrix is numerically well conditioned and hence invertible.

. A large variety of weighting functions exists here for trading lateral resolution for improved noise suppression (contrast), but one always ends up with a compromise between the two~\cite{Harris1978}.

@@

@@

Let $\X[n]$ be the data received by an array, pre-steered to all pixels in the image, and $\w$ be the window applied to these spatial data. Then we defined the beamformer output $z$ as the weighted sum of the data:
%
\begin{align}
z[\w,n] = \w\H\X[n] = \sumb{m=0}{M-1} w_m x_m^*[n]
\end{align}
%
From a set of $P$ windows the LCA beamformer selects the window $p$ that yield the least output power:
%
\begin{align}
\argmin{p}{E\big\{|z|^2\big\}} = \argmin{p}{E\big\{|\w_p\H\X[n]|^2\big\}} \qquad\text{subject to}\qquad \w\H\a = 1
\end{align}
%
We can estimate $E\big\{|\w_p\H\X[n]|^2\big\}$ as
%
\begin{align*}
\sigma^2_{z_p}[n] = \frac{1}{N_k} \sumb{n'=n-K}{n+K} | z_p[n'] |^2
\end{align*}
%
To add some extra flexibility, we can also estimate the variance by a weighted combination of local pixels
%
\begin{align*}
\sigma^2_{z_p}[x,n] = \frac{1}{N_x N_k} \sumb{x'=x-X}{x+X} \sumb{n=n-K}{n+K} w_\text{prf}[x',n']\big| z_p[x',n'] \big|^2
\end{align*}
%
where $w_\text{prf}[x',n']$ is a normalized 2 dimensional weight function.

Kaiser:
%
\begin{align*}
w[n] = \frac{I_0\big( \pi\alpha\sqrt{1-(\frac{2n}{N-1}-1)^2}\big)}{I_0(\pi\alpha)}
\end{align*}
%
where $I_0$ is a zeroth order modified Bessel function of the first kind.

@@

>>>>

The beamformer output $z[n]$ can now be expressed as the weighted sum of all the delayed data samples:
\begin{align}
z[n] = \w\H[n]\x[n] = \bmat{w_0[n]\\w_1[n]\\\vdots\\w_{M-1}[n]}^H \bmat{x_0[n]\\x_1[n]\\\vdots\\x_{M-1}[n]},\label{z}
\end{align}
where $w_m$ is the weight factor assigned to channel $m$. We will refer to a set of weights as a window, but it is also commonly called a taper function. Applying a window that trails off towards the edges improves noise suppression at the cost of reduced resolution~\cite{Harris1978}. Real windows lead to symmetric responses, while complex weights allow asymmetric responses.

@@

MVDR seeks to minimize the power of the beamformer, under the constraint of a unity gain in the desired direction $\a$~\cite{Capon1969}:
%
\begin{align}
\underset{\w[n]}{\argmin}\, E\{|z[n]|^2\} &= \underset{\w[n]}{\argmin}\, \w[n]\R[n]\w\H[n]\\
\text{subject to } \w[n]\a &= 1
\end{align}
%
where $\a$ is a steering vector and $\R[n] = E\{\x\x\H\}$ is the spatial covariance matrix. This is a convex optimization problem with the solution:
%
\begin{gather}
\vec w[n] = \frac{\Ri[n]\a}{\a\T\Ri[n]\a},\label{weights}
\end{gather}
where $\R=E\{\x[n]\x\H[n]\}\in\mathbb{C}^{M,M}$ is the spatial covariance matrix for the full array. The problem lies in estimating and inverting this spatial covariance matrix. Spatial averaging is needed to avoid signal cancellation~\cite{Kailath1985}, temporal averaging is needed to obtain true speckle statistics~\cite{Synnevag2009a}, and diagonal loading is needed to improve robustness to parameter errors~\cite{Cox1987,Maksym1979}. The steps are also needed to ensure that the covariance matrix is numerically well conditioned and hence invertible.

@@

From a set of $P$ windows the LCA beamformer selects the window $p$ that yields the least output power:
%
\begin{align}
\underset{p}{\argmin}\, E\{|z[n]|^2\} = \underset{p}{\argmin}\, E\big\{|\w_p\H\x[n]|^2\big\}\\
\qquad\text{subject to}\qquad \w\H\a = 1
\end{align}
%
We can estimate $E\big\{|\w_p\H\x[n]|^2\big\}$ as
%
\begin{align*}
\sigma^2_{z_p}[n] = \frac{1}{N_k} \sumb{n'=n-K}{n+K} | z_p[n'] |^2
\end{align*}
%
To add some extra flexibility, we can also estimate the variance by a weighted combination of local pixels
%
\begin{align*}
\sigma^2_{z_p}[x,n] = \frac{1}{N_x N_k} \sumb{x'=x-X}{x+X} \sumb{n=n-K}{n+K} w_\text{prf}[x',n']\big| z_p[x',n'] \big|^2
\end{align*}
%
where $w_\text{prf}[x',n']$ is a normalized 2 dimensional weight function.

@@

@@

\begin{align}
\w[\beta, \phi] = \text{kaiser}(\beta) = \bmat{
1 \\
e^{-j\frac{md}{\lambda}\sin(\phi)} \\
\vdots\\
e^{-j\frac{(M-1)d}{\lambda}\sin(\phi)}
}
\end{align}
